---
title: "‘Omics Project"
author: "Aeriel Belk, Amanda Walz, Rebecca Cheek, Kyle Root"
date: "11/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is an R-markdown we can use!!

Steps: 

1. import and early explore data
2. then, see what analyses we can do


Do some analysis of the metadata:

 - species
 - latitude/longitude
 - distribution over counties
 
```{r libraries}
library(seqinr)

```

``` {r metadata}
library(seqinr)
read.fasta()

```


genomic analysis

 - phyloseq?
 
```{r genomics}
library(phyloseq)

```


steps to read in:
1) list.files with full names option to get names of all files in folder, save in object (char vector)
2) pipe that vector into map function that does read.fasta (play around with just the first file first)
3) result should be a long list

seqinr -> read.fasta

```{r read-in}
#1) list files into vector
files <- list.files(path = 'consensus_sequences', full.names = TRUE)

#2) make map function to read in fasta files

read.fasta("consensus_sequences/W006_2013-07-10_USA_CA_SanDiego_-116.946_32.885.fa",
           as.string = TRUE,
           forceDNAtolower = FALSE)

fasta <- mapply(read.fasta())
?mapply()
```









